---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# P2C2M.SNAPP

P2C2M.SNAPP uses posterior predictive checks to identify violations to the multispecies coalescent model as implemented in the 'SNAPP' phylogeny estimation program. It was designed to be as accurate and user-friendly as possible.

## Installation

To install P2C2M.SNAPP, first download the gunzipped tarball and then install from source in R with:
```{r eval = FALSE}
install.packages("path/to/P2C2M.SNAPP_1.0.0.tar.gz", repos = NULL, type = "source")
```

**Dependencies:**  

*R*:  
R (>= 3.5.0),  
ape (>= 5.3),  
ggplot2 (>= 3.2.0),  
graphics (<= 3.5.0),  
grDevices (>= 3.5.0),  
gsubfn (>= 0.7),  
ggtree (>= 1.14.6) - not on CRAN (see below),  
KRIS (>= 1.1.1),  
stats (>= 3.5.0),  
utils (>= 3.5.0)  

The easiest way to install ggtree is with code from the Bioconductor website:
```{r eval = FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
  install.packages("BiocManager")

BiocManager::install("ggtree")
```

*Additional*:  
fastsimcoal2 (Excoffier et al., 2013: available at [http://cmpg.unibe.ch/software/fastsimcoal2/])
See fastsimcoal2 website for if you encounter errors - they may be due to the version of gcc on your computer.

*System*:
Mac users need to have XQuartz installed for fastsimcoal2 to run: available at [https://www.xquartz.org]


More installation information can be found in the tutorial or vignette.

## Use
The P2C2M.SNAPP package contains a single function, *run.p2c2m.snapp*. It requires as input the tree, log, and .xml files from a 'SNAPP' analysis as well as a simple metadata text file. With *run_mode = 1*, P2C2M.SNAPP will extract parameter values from the input files and simulate posterior predictive datasets. Users must then analyze these posterior predictive datasets with 'SNAPP'. We recommend performing these analyses with a computer cluster, as 'SNAPP' is quite computationally demanding. After analyzing the posterior predictive datasets with 'SNAPP', using *run_mode = 2* will calculate and compare summary statistics between the empirical and the posterior predictive datasets to identify possible model violations in the empirical dataset. A full tutorial is available from the P2C2M GitHub page [https://github.com/P2C2M].

## Citation
Please cite P2C2M.SNAPP as: Duckett, D.J., Pelletier, T.A., and Carstens, B.C. Submitted. Identifying model violations under the Multispecies Coalescent model with P2C2M.SNAPP.
